---
title: "Regression_Project_Final"
author: "Aaron Goletz, Alex Aubrey, Malini Senthilkumar, Angela Gaetano"
date: "December 7, 2017"
output:
  html_document: default
  word_document: default
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(scales)
library(stats)
library(MASS)
library(leaps)
library(car)
library(GGally)
library(corrplot)
library(knitr)
```

#1) An examination of the distributions of the variables

```{r, results='asis'}
comm_prop <- read.csv("comm_prop.csv", header=TRUE)
comm_prop$W2MiDT <- factor(comm_prop$W2MiDT, labels = c("No", "Yes"))

kable(head(comm_prop), format = "html", align='c', caption = "Example of Data") #Head function will return the FIRST 6 rows of the dataset

```

```{r}
summary(comm_prop) #Summary function returns a compact analysis of the dataset
```

**Observations**: The sample Data and Summary provide a sense of the data. 
Just by looking at the brief summary, we observe that a few variables like SqFt and Taxes have a higher mean than their medians. This indicates the variables are probably right-skewed. We will explore the distributions further with histogram plots.

**Variables are : **  
    RentRate  - Continuous Variable   
    Age       - Continuous Variable   
    OperExp   - Continuous Variable  
    VacRate   - Continuous Variable   
    SqFt      - Continuous Variable     
    Taxes     - Continuous Variable 
    W2MiDT    - Categorical Variable   
    
    
## Histograms for distribution visualization
```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
plot(comm_prop$W2MiDT , xlab="Within 2 Miles of Downtown", ylab="Frequency")
boxplot(RentRate~W2MiDT, data=comm_prop)
```

**Observations**: Histograms help us better understand the distribution of continuous variables. 
Above, we have plotted all continuous variables into histograms and have created a bar plot and boxplot for categorical variables.

**We noticed the following:**  
Rental Rate         - looks normally distributed  
Age                 - appears bimodal. Building construction occurred more frequently about 15 years ago and within the last few years.  
OperExp             - contains a slight amount of left/negative skewness  
VacRate             - contains strong Right/positive skewness  
SqFt                - contain a slight amount of right skewness  
Taxes               - contain a slight amount of right skewness  
W2MiDTNo            - has a higher frequency  
*
The boxplot indicates that the rental rates are higher when the building is within two miles of downtown.   

#2)  An examination of the correlations for all continuous variables


**Correlation Matrix with scatterplots and correlation coefficients**
```{r,results="hide",fig.keep='all',message=FALSE}
ggpairs(comm_prop[,1:7] , 
        lower = list(continuous = wrap("points", color = "red", alpha = 0.5)), 
        diag = list(continuous = wrap("densityDiag",  color = "blue", alpha = 0.5)))

```


**Observations:**  
SqFt and Taxes        - have a perfect correlation of 1.00.  
RentRate and OperExp  - have a moderate positive correlation.   
RentRate and SqFt     - have a strong positive correlation.    
RentRate and Taxes    - share the same correlation as of .584.   

**Other correlations to note are:**
VacRate and OperExp   - a negative correlation of -.366   
Age and OperExp       - a very small correlation   
Age and RentRate      - a very small negative correlation  


#3)  The identification and evaluation of a suitable regression model for predicting rental rates


##a) Manual Iteration

### Fit Full Model
```{r, warning=F}
m <- lm(RentRate ~ ., data=comm_prop)
summary(m)
```

**Observations:**
Results show the model's significant variables are Age, OperExp, and W2MiDT.  The Adjusted R-squared value is 0.73, with an F-statistic of 45.61 and a highly significant p-value of <2.2e-16.  The model will now be trimmed to remove non-significant variables.

First, VacRate will be dropped due to non-significance (p-value 0.408). In addition, VacRate is unavailable for use as a final predictor due to unavailable data. Also, the variable Taxes will be dropped from the model due to multicolinearity with SqFt (correlation coefficient 1.0) as evidenced by the correlation matrix above. Since Taxes is a result of SqFt, we are dropping Taxes and retained SqFt in our model.


```{r, warning=F}
m.1 <- lm(RentRate ~ Age + OperExp + SqFt + W2MiDT, data=comm_prop)
summary(m.1)
```

**Observations**: All variables are highly significant.  Model m.1's summary shows an increased Adjusted R-squared value of .7313, a slightly increased F-Statistic of 68.35, and no notable change in the p-value.


##b) Stepwise Iterations

```{r,warning=FALSE,results="hide"}
comm_prop.full <- lm(RentRate ~ .*., data = comm_prop)
stepAIC(comm_prop.full, direction ="backward")
```

The stepAIC yields the following formula for an optimal linear model:

    formula = RentRate ~ Age + OperExp + VacRate + Taxes + W2MiDT + 
    Age:OperExp + Age:VacRate + OperExp:VacRate + OperExp:W2MiDT + 
    Taxes:W2MiDT


```{r, warning=F}
m.step <- lm(formula = RentRate ~ Age + OperExp + VacRate + Taxes + W2MiDT + 
    Age:OperExp + Age:VacRate + OperExp:VacRate + OperExp:W2MiDT + 
    Taxes:W2MiDT, data = comm_prop)
summary(m.step)
```    

Removing VacRate and associating interactions.

```{r, warning=F}
m.step <- lm(formula = RentRate ~ Age + OperExp + Taxes + W2MiDT + 
    Age:OperExp + OperExp:W2MiDT + Taxes:W2MiDT, data = comm_prop)
summary(m.step)
```    

The m.step model has an Adjusted R-squared value of 0.8171 and an F-Statistic of 64.18 with a highly significant p-value, <2.2e-16.

*The Taxes variable is not significant, but it will stay in the model as its interaction with W2MiDT is highly significant.*


##c) All Subsets Regression (ASR) Iteration 

```{r}
subsets <- regsubsets(RentRate ~ .*., data=comm_prop)
plot(subsets, scale="adjr2")

```

Based on the subsets plot, our variable selection should be Age, OperExp, W2MiDT, Age:OperExp, Age:VacRate, Age:Taxes, OperExp:W2MiDT, and SqFt:W2MiDT.

No VacRate interaction will be used, as it is an unknown variable in our final prediction.

```{r}
m.sub <- lm(RentRate ~ Age+OperExp+W2MiDT+Age:OperExp+Age:Taxes+OperExp:W2MiDT+SqFt:W2MiDT, data=comm_prop)
summary(m.sub)

```

**The results of model m.sub are an Adjusted R-squared value of 0.8175, an F-statistic of 56.45, and a highly significant p-value of <2.2e-16.**

The subset Age:Taxes is not significant, so it can be trimmed from the model.  

```{r, warning=F}
m.sub1 <- lm(RentRate ~ Age+OperExp+W2MiDT+Age:OperExp+OperExp:W2MiDT+SqFt:W2MiDT, data=comm_prop)
summary(m.sub1)
```


**The Adjusted R-squared value decreased slightly to 0.8171.  The F-Statistic increased to 64.18 and the p-value had no change.**


## d) Review and Interpretation of diagnostic plots
              
                  **Adjusted R-sqrd   F-Statistic**
              m.1       0.7313            68.35 
              m.step    0.8171            64.18
              m.sub1    0.8171            64.18
      

**Models m.step and m.sub1 appear to to have the same values for F-Statistic and Adjusted R-squared and therefore are more appropriate for a predictive model.  Further examination will need to be conducted to determine the most appropriate model.**

###Leverage Plot
```{r}
leveragePlots(m.1,pch=16)

leveragePlots(m.step,pch=16)

leveragePlots(m.sub1,pch=16)
```

**Observations**: Since the leverage plots for m.1 have slopes farther from y=0, m.1 seems to have the most significant variables compared to the other models. 


###Residuals Variance Plot
```{r}
library(ggplot2)
m_df <- fortify(m.1)
ggplot(m_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.1")
m.step_df <- fortify(m.step)
ggplot(m.step_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.step1")
m.sub_df <- fortify(m.sub1)
ggplot(m.sub_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.sub1")
```

**Observations**: Models m.step and m.sub1 both seem to cluster in the middle section of the plot.  This suggests non-constant variance.


###Checking for constant variance###
```{r}
car::ncvTest(m.1)
car::ncvTest(m.step)
car::ncvTest(m.sub1)
```

**Conclusion**: Significantly low p-values for the ncvTest confirm that both the m.step and m.sub1 models have non-constant variance.  The m.1 model maintains constant variance, however.



###Visual Representation of Normality of Residuals###
```{r}
par(mfrow=c(2,3))

hist(m.1$residuals)
hist(m.step$residuals)
hist(m.sub1$residuals)

qqnorm(m.1$residuals, pch=16)
qqline(m.1$residuals)

qqnorm(m.step$residuals, pch=16)
qqline(m.step$residuals)

qqnorm(m.sub1$residuals, pch=16)
qqline(m.sub1$residuals)
```

**Observations**: The residuals appear to be approximately normal.


###Testing Normality of Residuals###
```{r}
shapiro.test(m.1$residuals)
shapiro.test(m.step$residuals)
shapiro.test(m.sub1$residuals)
```

**Conclusion**: Significantly low p-values for the Shapiro-Wilk test shows that both the m.step and m.sub1 models are not normally distributed.  The m.1 model has approximately normal residuals.


###Test for Multicolinearity
```{r}
round(vif(m.1),3)
round(vif(m.step),3)
round(vif(m.sub1),3)

```
**Conclusion**:  Tests suggest no Multicolinearity.


# 4) Application of the Model

Without further data transformation, due to the non-constant variance and disnormality of the m.step and m.sub1 models, **m.1 is the most appropriate model to use in our prediction.**

**73% of the RentRate's Variation can be explained by Age, OperExp, SqFt and W2MiDT**

###Application using mathematical function  

$$RentRate = 12.05 - 0.104Age + 0.237OperExp + .00000664SqFt + 1.435W2MiDTYes + e$$

```{r,ECHO=FALSE}
12.05 - 0.104*9 + 0.237*13 + .00000664*40 + 1.435*0

```

###Application using predict() function (provides interval)
```{r}
new_obs = data.frame(Age=9, SqFt=40,OperExp=13,W2MiDT="No")
predict(m.1, newdata = new_obs, interval="predict")

```

####**Prediction: We are 95% confidence that the value of RentRate will be between $12,160.37 and $16,235.17.  With the combined costs of Operating Expenses and Taxes, $13,540.00, we cannot say with confidence that we will be able to "offset the costs of ownership."  We should sell the building.**

***

###Further Discussions  

Other data points that would potential help the ability to predict Rental Rate are:  
1) More details about the costs that are incorporated in operating expenses so we can consider missing operation costs such as insurance, security, grounds-keeping, etc.    
2) More detailed location  
3) Market dynamics and trends  
4) The domain of the market or consumers  
