---
title: "Exploring the Prediction of Rent Rate using Neighboring Properties Data"
author: "Group 1"
date: "November 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) An examination of the distributions of the variables
```{r}
comm_prop <- read.csv("comm_prop.csv")
head(comm_prop,5)
```



```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
hist(comm_prop$W2MiDT)

```


## 2) An examination of the correlations for all of the continuous variables.
```{r comm-prop.2, echo=FALSE}
cor(comm_prop[1:6])
```

####  The Correlation Matrix Scatterplot of continuous variables.

There are moderate relationships between <strong>OperExp</strong> and <strong>RentRate</strong>, <strong>OperExp</strong> and <strong>SqFt</strong>, <strong>OperExp</strong> and <strong>Taxes</strong>.

There are strong relationships between <strong>RentRate</strong> and <strong>SqFt</strong>, <strong>RentRate</strong> and 
<strong>Taxes</strong>, and a very strong relationship between <strong>Taxes</strong> and <strong>SqFt</strong>.

```{r comm_prop.2A, echO=FALSE}
pairs(comm_prop[1:6], pch=16)
```
