---
title: "Exploring the Prediction of Rent Rate using Neighboring Properties Data"
author: "Group 1"
date: "November 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(Hmisc)
library(scales)
library(stats)
library(car)
library(GGally)
library(corrplot)
```

## 1) An examination of the distributions of the variables

```{r}
comm_prop <- read.csv("comm_prop.csv", header=TRUE)
attach(comm_prop)
head(comm_prop,5)
```



```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
hist(comm_prop$W2MiDT)

```


## 2) An examination of the correlations for all of the continuous variables.

```{r comm-prop.2, echo=FALSE}
cor(comm_prop[1:6])
```


```{r comm-prop.2, echo=FALSE}
corrplot(cor(comm_prop[,1:6]), method="ellipse")
```


```{r comm_prop.2A, echO=FALSE}
 ggpairs(comm_prop[,1:6] , 
        lower = list(continuous = wrap("points", color = "red", alpha = 0.5)), 
        diag = list(continuous = wrap("densityDiag",  color = "blue", alpha = 0.5)))
```

The graph shows that there is a strong correlation between Taxes and SQft. 
Vacancy rate and operating expense shows negative correlation, which states that operating expense is close to zero if the building is available for rent. 

With this given sample data, it seems the profit will depend on how much rental rate is for the building. 
Concluding, rental rate will be our outcome variable.


```{r}
model_1 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + Age:SqFt + Age:W2MiDT + OperExp:VacRate + OperExp:SqFt + OperExp:W2MiDT + VacRate:SqFt + SqFt:W2MiDT, data=comm_prop)
summary(model_1)
```

```{r}
#remove VacRate:SqFt
model_2 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + Age:SqFt + Age:W2MiDT + OperExp:VacRate + OperExp:SqFt + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_2)
```

```{r}
#remove OpenExp:SqFt
model_3 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + Age:SqFt + Age:W2MiDT + OperExp:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_3)
```

```{r}
#remove Age:SqFt
model_4 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + Age:W2MiDT + OperExp:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_4)
```

```{r}
#remove Age:W2MiDT
model_5 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + OperExp:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_5)
```

```{r}
#remove OpenExp:VacRate
model_6 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:OperExp + Age:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_6)
```

```{r}
#remove Age:OpenExp  - Taking this as Model
model_7 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT + Age:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(model_7)


leveragePlots(model_7,pch=16)


```
```{r}
model_7_df <- fortify(model_7)

ggplot(model_7_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals")
```


**Checking for constant variance**
```{r}
car::ncvTest(model_7)

```

**Conclusion: Variance is constant**  

**Checking for Normality of Residuals**
```{r}
shapiro.test(model_7$residuals)

```

**The residuals are approximately normal**

####Visual Representation of Normality and Constant Variance Tests
```{r}
par(mfrow=c(2,2))
hist(model_7$residuals)
plot(model_7$residuals ~ comm_prop$RentRate,pch=16, main="Residual Plot")
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
qqnorm(model_7$residuals, pch=16)
qqline(model_7$residuals)
```

Tails are a bit out, which could suggest non-constant variance. However, we formally tested via the
ncvTest and did not reject the null hypothesis.

**Test for Multicolinearity**
```{r}
car::vif(model_7)

```
