---
title: "Exploring the Prediction of Rent Rate using Neighboring Properties Data"
author: "Group 1"
date: "November 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(scales)
library(stats)
library(car)
library(GGally)
library(corrplot)
```

## 1) An examination of the distributions of the variables
```{r}
comm_prop <- read.csv("comm_prop.csv")
#comm_prop$W2MiDT <- factor(comm_prop$W2MiDT, labels = c("No", "Yes"))
head(comm_prop,5)
```



```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
hist(comm_prop$W2MiDT)

```

VacRate's distribution contains strong Right/positive skewness. SqFt contains a slight amount of 
right skewness. OperExp contains a slight amount of left/negative skewness. Age has an unusual 
ditribution shape to it as well.


```{r}
round(cor(comm_prop),2)
```

```{r}
library(GGally)
ggpairs(comm_prop[,1:6] , 
        lower = list(continuous = wrap("points", color = "red", alpha = 0.5)), 
        diag = list(continuous = wrap("densityDiag",  color = "blue", alpha = 0.5)))
```

```{r}
library(corrplot)
corrplot(cor(comm_prop[,c(1:6)]), method="ellipse")
```

RentRate and OperExp have a moderate correlation. RentRate and SqFt have a moderate correlation as well.
RentRate and Taxes share the same correlation with .58. W2MiDT has the strongest 
correlation with RentRate with a correlation coefficent of.70. SqFt and Taxes have a perfect correlation
with 1.00.


```{r,warning=FALSE}
library(leaps)
subsets <- regsubsets(RentRate ~ ., data=comm_prop)
plot(subsets, scale="adjr2")

```

####^^ Based on the plot, our variable selection should be Age, OperExp, Taxes, and W2MiDT.

```{r}
m <- lm(RentRate ~ Age+OperExp+W2MiDT+Taxes, data=comm_prop) # + Age:VacRate + OperExp:W2MiDT + SqFt:W2MiDT, data=comm_prop)
summary(m) #added interactive terms from Malini's result and got potentially more accurate model

```
#Review and Interpretation of diagnostic plots
```{r}
leveragePlots(m,pch=16)
```

```{r}
m_df <- fortify(m)

ggplot(m_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals")
```

**Checking for constant variance**
```{r}
car::ncvTest(m)

```

**Conclusion: Variance is constant**  

**Checking for Normality of Residuals**
```{r}
shapiro.test(m$residuals)

```

**The residuals are approximately normal**

####Visual Representation of Normality and Constant Variance Tests
```{r}
par(mfrow=c(2,2))
hist(m$residuals)
plot(m$residuals ~ comm_prop$RentRate,pch=16, main="Residual Plot")
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
qqPlot(m$residuals, pch=16)
# qqline(m$residuals)
```

Tails are a bit out, which could suggest non-constant variance. However, we formally tested via the
ncvTest and did not reject the null hypothesis.

**Test for Multicolinearity**
```{r}
car::vif(m)

```

**Results suggest no Multicolinearity**

Start on Predict: We are 95% confidence that the value of RentRate will be between 
14.17338 and 18.20759.
```{r}
new_obs = data.frame(Age=9, SqFt=40,OperExp=13,Taxes=5.40,W2MiDT=0)
predict(m, newdata = new_obs, interval="predict")
```


**Stepwise regression by manually removing variables **


