---
title: "Exploring the Prediction of Rent Rate using Neighboring Properties Data"
author: "Group 1"
date: "November 27, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(Hmisc)
library(scales)
library(stats)
library(car)
library(GGally)
library(corrplot)
```

## 1) An examination of the distributions of the variables

```{r}
comm_prop <- read.csv("comm_prop.csv", header=TRUE)
attach(comm_prop)
head(comm_prop,5)
```



```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
hist(comm_prop$W2MiDT)

```


## 2) An examination of the correlations for all of the continuous variables.

```{r comm-prop.2, echo=FALSE}
cor(comm_prop[1:6])
```


```{r comm-prop.2, echo=FALSE}
corrplot(cor(comm_prop[,1:6]), method="ellipse")
```


```{r comm_prop.2A, echO=FALSE}
 ggpairs(comm_prop[,1:6] , 
        lower = list(continuous = wrap("points", color = "red", alpha = 0.5)), 
        diag = list(continuous = wrap("densityDiag",  color = "blue", alpha = 0.5)))
```

The graph shows that there is a strong correlation between Taxes and SQft. 
Vacancy rate and operating expense shows negative correlation, which states that operating expense is close to zero if the building is available for rent. 

With this given sample data, it seems the profit will depend on how much rental rate is for the building. 
Concluding, rental rate will be our outcome variable.
