---
title: "Regression_Project_Final"
author: "Aaron Goletz"
date: "December 6, 2017"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(Hmisc)
library(scales)
library(stats)
library(MASS)
library(leaps)
library(car)
library(GGally)
library(corrplot)
library(knitr)
```

# 1) An examination of the distributions of the variables

```{r, results='asis'}
# comm_prop <- read.csv("C:/Users/aaron.goletz/Desktop/My Docs/Lipscomb Documents/Stats Analysis and Decision Modeling/MSDS 5043/comm_prop.csv")
comm_prop <- read.csv("comm_prop.csv", header=TRUE)
comm_prop$W2MiDT <- factor(comm_prop$W2MiDT, labels = c("No", "Yes"))

kable(head(comm_prop), format = "html", align='c', caption = "Example of Data") #Head function will return the FIRST 6 rows of the dataset

```

```{r}
# tail(comm_prop) #Tail function will return the LAST 6 rows of the dataset
summary(comm_prop) #Summary function returns a compact analysis of the dataset
```

Observations: Head, Tail and Summary metrics gives sense of the data. 
Just by looking at the brief summary, we can notice few variables like SqFt, Taxes has a higher mean than its median. This indicates the variables are probably right skewed. We will explore further with histogram plots.

Variables are :   
    RentRate  - Continuous Variable   
    Age       - Continuous Variable   
    OperExp   - Continuous Variable  
    VacRate   - Continuous Variable   
    SqFt      - Continuous Variable     
    Taxes     - Continuous Variable 
    W2MiDT    - Categorical Variable   
    
    
###Histograms for distribution visualization
```{r comm_prop, echo=FALSE}
par(mfrow = c(3,3))
hist(comm_prop$RentRate)
hist(comm_prop$Age)
hist(comm_prop$OperExp)
hist(comm_prop$VacRate)
hist(comm_prop$SqFt)
hist(comm_prop$Taxes)
plot(comm_prop$W2MiDT , xlab="Within 2 Miles of Downtown", ylab="Frequency")
boxplot(RentRate~W2MiDT, data=comm_prop)
```

Observations: Histograms help us better understand the distribution of continuous variables. 
Above, we have plotted all continuous variables into histograms and bar plot and boxplot for Categorical Variable.

We noticed the following:
Rental Rate         - looks normally distributed
Age                 - appears bimodal. It has an unusual distribution shape to it. It looks like there were buildings built recently and also 
                      around 15 years ago it was higher.
OperExp             - contains a slight amount of left/negative skewness.
VacRate             - contains strong Right/positive skewness
SqFt                - contain a slight amount of right skewness
Taxes               - contain a slight amount of right skewness
W2MiDTNo            - has a higher frequency.


#2)  An examination of the correlations for all continuous variables

**Correlation Matrix**
```{r}
round(cor(comm_prop[1:6]),3)
```

**Correlation Matrix with scatterplots and correlation coefficients**
```{r,results="hide",fig.keep='all',message=FALSE}
ggpairs(comm_prop[,1:7] , 
        lower = list(continuous = wrap("points", color = "red", alpha = 0.5)), 
        diag = list(continuous = wrap("densityDiag",  color = "blue", alpha = 0.5)))

# 'ggpairs' function is derived from 'GGally' library and draws a martix of plots of the diven dataset
```

**Correlation Matrix with ellipse representation of scatterplots**
```{r}
corrplot(cor(comm_prop[,c(1:6)]), method="ellipse")
#'corrplot' function is derived from the 'corrplot' library and returns the correlations of multiple factors of the dataset in graphical table
```

Observations:  
SqFt and Taxes        - have a perfect correlation of 1.00.
RentRate and OperExp  - have a moderate positive correlation. 
RentRate and SqFt     - have a strong positive correlation.  
RentRate and Taxes    - share the same correlation as of .584. 

Other correlations to note are:
VacRate and OperExp   - a negative correlation of -.366 
Age and OperExp       - a very small correlation 
Age and RentRate      - a very small negative correlation


#3)  The identification and evaluation of a suitable regression model for predicting rental rates

A manual iteration selection method will be used first.  
 
Manual Iteration is a process of building the best-fit model by manually observing the data set and eliminating the factors without a significant p-value one by one. Then observing the new model’s adjusted R-sqaure value.
The process of elimination is done for every variable that has the highest p-value (i.e is LEAST significant).

For this, we use the ‘lm’ function, to fit Linear Models to the datasets and execute the ‘summary’ function on the model. 

The results will later be compared to results from Stepwise and All Subsets Regression (ASR) selection methods in order to select the most appropriate linear model.


##a) FIT FULL MODEL

```{r, warning=F}
m <- lm(RentRate ~ ., data=comm_prop)
summary(m)
```

Observations:
Results show the model's significant variables are Age, OperExp, and W2MiDT.  The Adjusted R-squared value is good at 0.73, with an F-statistic of 45.61 and a high significant p-value of <2.2e-16.  The model will now be trimmed to remove non-significant variables.

First, VacRate will be dropped due to non-significance in addition to it being an unknown variable in our final prediction. VacRate data is not as readily available as the others. Also.,the variable Taxes (0.429) will be dropped from the model, due to multicolinearity with SqFt (0.434).  



## ITERATE MANUALLY

```{r, warning=F}
m.4 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt + W2MiDT, data=comm_prop) # drop Taxes
summary(m.4)

m.3 <- lm(RentRate ~ Age + OperExp + VacRate + SqFt, data=comm_prop) # drop W2MiDT
summary(m.3)

m.2 <- lm(RentRate ~ Age + OperExp + VacRate + W2MiDT, data=comm_prop) # drop SqFt
summary(m.2)

m.1 <- lm(RentRate ~ Age + OperExp + SqFt + W2MiDT, data=comm_prop) # drop VacRate
summary(m.1)
```

Observations: All variables are highly significant.  Model m.1's summary shows an increased Adjusted R-squared value of .7313, an increased F-Statistic of 68.35, and no notable change in the p-value.*


##b) Stepwise Iterations

```{r,warning=FALSE,results="hide"}
comm_prop.full <- lm(RentRate ~ .*., data = comm_prop)
stepAIC(comm_prop.full, direction ="backward")
```

The stepAIC yields the following formula for an optimal linear model:

    formula = RentRate ~ Age + OperExp + VacRate + Taxes + W2MiDT + 
    Age:OperExp + Age:VacRate + OperExp:VacRate + OperExp:W2MiDT + 
    Taxes:W2MiDT


```{r, warning=F}
m.step <- lm(formula = RentRate ~ Age + OperExp + VacRate + Taxes + W2MiDT + 
    Age:OperExp + Age:VacRate + OperExp:VacRate + OperExp:W2MiDT + 
    Taxes:W2MiDT, data = comm_prop)
summary(m.step)
```    

Removing VacRate and associating interactions.

```{r, warning=F}
m.step <- lm(formula = RentRate ~ Age + OperExp + Taxes + W2MiDT + 
    Age:OperExp + OperExp:W2MiDT + Taxes:W2MiDT, data = comm_prop)
summary(m.step)
```    

The m.step model has an Adjusted R-squared value of 0.8171 and an F-Statistic of 64.18 with a highly significant p-value, <2.2e-16.

*The Taxes variable is not significant, but it will stay in the model as its interaction with W2MiDT is highly significant.*


##c) ASR Iteration

```{r}
subsets <- regsubsets(RentRate ~ ., data=comm_prop)
plot(subsets, scale="adjr2")

```

Based on the subsets plot, our variable selection should be Age, OperExp, W2MiDT, Age:OperExp, Age:VacRate, Age:Taxes, OperExp:W2MiDT, and SqFt:W2MiDT.

No VacRate interaction will be used, as it is an unknown variable in our final prediction.

```{r}
m.sub <- lm(RentRate ~ Age+OperExp+W2MiDT+Age:OperExp+Age:Taxes+OperExp:W2MiDT+SqFt:W2MiDT, data=comm_prop)
summary(m.sub)

```

**The results of model m.sub are an Adjusted R-squared value of 0.8175, an F-statistic of 56.45, and a highly significant p-value of <2.2e-16.**

The subset Age:Taxes is not significant, so it can be trimmed from the model.  

```{r, warning=F}
m.sub1 <- lm(RentRate ~ Age+OperExp+W2MiDT+Age:OperExp+OperExp:W2MiDT+SqFt:W2MiDT, data=comm_prop)
summary(m.sub1)
```


**The Adjusted R-squared value decreased slightly to 0.8171.  The F-Statistic increased to 64.18 and the p-value had no change.**


##Review and Interpretation of diagnostic plots
              
                <strong>Adjusted R-sqrd   F-Statistic</strong> 
              m.2       0.7313            68.35 
              m.step    0.8171            64.18
              m.sub1    0.8171            64.18
      

###Models m.step1 and m.sub1 appear to to have the same values for F-Statistic and Adjusted R-squared the most appropriate predictive model.  Further examination will need to be conducted to determine the most appropriate model.

```{r}
leveragePlots(m.1,pch=16)

leveragePlots(m.step,pch=16)

leveragePlots(m.sub1,pch=16)
```

Observations: Leverage Plots are a way to visualize which variables are significant.
– If the red line is not close to the horizontal line at y = 0, the variable is considered to be significant


###Residuals Variance Plot
```{r}
library(ggplot2)
m_df <- fortify(m.1)
ggplot(m_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.1")
m.step_df <- fortify(m.step)
ggplot(m.step_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.step1")
m.sub_df <- fortify(m.sub1)
ggplot(m.sub_df, aes(x=.fitted, y=.resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = 2) +
  labs(x="Fitted Values", y="Residuals", title="m.sub1")
```

Observations: Models m.step and m.sub1 both seem to cluster in the middle section of the plot.  This suggests non-constant variance.


###Checking for constant variance###
```{r}
car::ncvTest(m.1)
car::ncvTest(m.step)
car::ncvTest(m.sub1)
```

Conclusion: Significantly low p-values for the ncvTest confirm that both the m.step and m.sub1 models have non-constant variance.  The m.1 model maintains constant variance, however.


###Checking for Normality of Residuals###
```{r}
shapiro.test(m.1$residuals)
shapiro.test(m.step$residuals)
shapiro.test(m.sub1$residuals)
```

Conclusion: Significantly low p-values for the Shapiro-Wilk test confirm that both the m.step and m.sub1 models' are not normally distributed.  The m.1 model is normal.

###Visual Representation of Normality and Constant Variance Tests###
```{r}
par(mfrow=c(3,3))

hist(m.1$residuals)
plot(m.1$residuals ~ comm_prop$RentRate,pch=16, main="Residual Plot")
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
qqnorm(m.1$residuals, pch=16)
qqline(m.1$residuals)

hist(m.step$residuals)
plot(m.step$residuals ~ comm_prop$RentRate,pch=16, main="Residual Plot")
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
qqnorm(m.step$residuals, pch=16)
qqline(m.step$residuals)

hist(m.sub1$residuals)
plot(m.sub1$residuals ~ comm_prop$RentRate,pch=16, main="Residual Plot")
abline(h = 0, lty = 3)  # adds a horizontal dashed line at y = 0
qqnorm(m.sub1$residuals, pch=16)
qqline(m.sub1$residuals)
```


###Test for Multicolinearity###
```{r}
car::vif(m.1)
car::vif(m.step)
car::vif(m.sub1)

```
Conclusion:  Tests suggest no Multicolinearity.


**Without further data transformation, due to the non-constant variance and disnormality of the m.step and m.sub1 models, m.1 is the most appropriate model to use in our prediction.**

```{r}
new_obs = data.frame(Age=9, SqFt=40,OperExp=13,W2MiDT="No")
predict(m.1, newdata = new_obs, interval="predict")

new_obs = data.frame(Age=9, SqFt=40,OperExp=13,Taxes=.54,W2MiDT="No")
predict(m.step, newdata = new_obs, interval="predict")

new_obs = data.frame(Age=9, SqFt=40,OperExp=13,Taxes=.54,W2MiDT="No")
predict(m.sub1, newdata = new_obs, interval="predict")
```

**Prediction: We are 95% confidence that the value of RentRate will be between $12,120.37 and $16,235.17.  With the combined costs of Operating Expenses and Taxes, $13540.00, we cannot say with confidence that we will be able to "offset the costs of ownership."  We should sell the building.**
